---
title: "Maumee Community Analysis"
author: "Jordyn Stoll"
date: "April 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
maumee <- read.csv("~/Documents/GitHub/algalanalysis/algalcommunityanalysis/Community Analysis/Maumee/maumee.csv")
library(vegan)
str(maumee)
summary(maumee)
```

Use the Vegan package

Create shannon diversity, richness, and rare species
```{r}
speciesmatrix<- maumee [4:61]
maumee$shannon<- diversity(speciesmatrix, index = "shannon")
maumee$richness<- specnumber(speciesmatrix)
maumee$rarespecies<-rarefy(speciesmatrix, sample=min(maumee$Total.Counted))
```


Generalize Linear model
#add an offset value to consider sample effect into the Shannon model. Slope=1. 
```{r}
shannon.model<-glm(shannon~Treatment+offset(Total.Counted), data=maumee)
shannon.model
summary(shannon.model)
anova.shannon.model<-aov(shannon.model)
anova.shannon.model
summary(anova.shannon.model)
TukeyHSD(anova.shannon.model)
```

Plot the summary statistics for shannon diversity and species richness
#When using ggplot, use + between different layers
#Standard Error (se) is equal to sd/sqrt(length)
```{r}
library(plyr)
summarystatsShannonRichness<- ddply(maumee, "Treatment", summarize,
                     shannon.mean=mean(shannon), shannon.sd=sd(shannon),
                     shannon.se=sd(shannon)/sqrt(length(shannon)),richness.mean=mean(richness),
                      richness.sd=sd(richness), richness.se=sd(richness)/sqrt(length(richness)))

library(ggplot2)
ShannonDiversitysummaryplot<-ggplot(summarystatsShannonRichness, aes(Treatment, shannon.mean, fill=Treatment))+ 
  geom_bar(stat='identity',color='black')+
  geom_errorbar(aes(ymin=shannon.mean-shannon.sd, ymax=shannon.mean+shannon.sd))+
  theme(axis.text.x=element_text(angle = 45, hjust=1))+
  ggtitle("Maumee Shannon Diversity vs. Treatment")

ShannonDiversitysummaryplot

richnesssummaryplot<-ggplot(summarystatsShannonRichness, aes(Treatment,richness.mean, fill=Treatment))+ 
  geom_bar(stat='identity',color='black')+geom_errorbar(aes(ymin=richness.mean-richness.sd, ymax=richness.mean+richness.sd))+
    theme(axis.text.x=element_text(angle = 45, hjust=1))+
  ggtitle("Maumee Richness vs. Treatment")
richnesssummaryplot

```


NMDS- how different are communities between treatments? 
#fussy about low numbers. Good at deciphering similar communities.
#create matrices for the environment and for the species- opposites
#autotransform is a double sqrt(Wisconsin transformation)- Bray Curtis dissimilarity to decide similarity between observations- regress between points within samples to determine dismilarity- apply transformation then does pairwise comparison to determine similarity metric
#ordination rules of thumb 2-D- stess should be less than 0.2... if it is over 0.2 you need to fiddle with the transformation until it works.
#environm gives centroids, while adonic gives stats
#perm is the number of permuations.. or times it tries to fit
```{r}
speciesm<-maumee[2:37,4:61]
environm<-maumee[2:37,1:3]
speciesord<-metaMDS(speciesm,
             autotransform = T)
speciesord
environfit<-envfit(speciesord~Treatment, environm, perm=999)
summary(environfit)
environfit
betad<- betadiver(speciesm, "z")
adonis(betad~Treatment, environm, perm=999)
#since this is significant, we need to do a pairwise comparison of each treatment- subset treatments and comapre
#add in pal to color code treatments.
plot(speciesord, disp='species', type='n', bg=pal[Treatment])
with(environm, points(speciesord, display='species', col='black'))
plot(environfit, col='green')

```

