---
title: "Maumee Community Analysis"
author: "Jordyn Stoll"
date: "April 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
maumee <- read.csv("~/Desktop/maumee.csv")
library(vegan)
str(maumee)
summary(maumee)
```

Use the Vegan package

NDMS- how different are communities between treatments? 

Create shannon diversity, richness, and rare species
```{r}
speciesmatrix<- maumee [4:61]
maumee$shannon<- diversity(speciesmatrix, index = "shannon")
maumee$richness<- specnumber(speciesmatrix)
maumee$rarespecies<-rarefy(speciesmatrix, sample=min(maumee$Total.Counted))
```


Generalize Linear model
#add an offset value to consider sample effect into the Shannon model. Slope=1. 
```{r}
shannon.model<-glm(shannon~Treatment+offset(Total.Counted), data=maumee)
shannon.model
summary(shannon.model)
anova.shannon.model<-aov(shannon.model)
anova.shannon.model
summary(anova.shannon.model)
TukeyHSD(anova.shannon.model)
```

Plot the summary statistics for shannon diversity and species richness
#When using ggplot, use + between different layers
#Standard Error (se) is equal to sd/sqrt(length)
```{r}
library(plyr)
summarystatsShannonRichness<- ddply(maumee, "Treatment", summarize,
                     shannon.mean=mean(shannon), shannon.sd=sd(shannon),
                     shannon.se=sd(shannon)/sqrt(length(shannon)),richness.mean=mean(richness),
                      richness.sd=sd(richness), richness.se=sd(richness)/sqrt(length(richness)))
library(ggplot2)
ShannonDiversitysummaryplot<-ggplot(summarystatsShannonRichness, aes(Treatment, shannon.mean, fill=Treatment))+ geom_bar(stat='identity',color='black')+geom_errorbar(aes(ymin=shannon.mean-shannon.sd, ymax=shannon.mean+shannon.sd))
ShannonDiversitysummaryplot

richnesssummaryplot<-ggplot(summarystatsShannonRichness, aes(Treatment,richness.mean, fill=Treatment))+  geom_bar(stat='identity',color='black')+geom_errorbar(aes(ymin=richness.mean-richness.sd, ymax=richness.mean+richness.sd))
richnesssummaryplot

```

