---
title: "USGS HABs 2017"
author: "Dave Costello"
date: "8/23/2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~Documents/GitHub/algalanalysis/algalcommunityanalysis/Community Analysis/")
HAB2017<- read.csv("~/Documents/GitHub/algalanalysis/Daves code/HAB2017_incub.csv")

library(lme4)
library(MuMIn)
```

##Function for calculating log response ratios from incubator experiment
Function calculates log RR and the standard error of RR for each treatment (treatment/control) excluding initial data. Returns a dataframe with treatment, RR and se as columns.

```{r}
logRR <- function(HAB2017){
  HAB2017$Treatment <- HAB2017$Treatment[drop=T]
  trt <- levels(HAB2017$Treatment)
  trt.no <- length(trt)-1
  means <- with(HAB2017,tapply(Chla,Treatment,mean,na.rm=T))
  sds <- with(dataset,tapply(Chla,Treatment,sd,na.rm=T))
  ns <- with(dataset,tapply(Chla,Treatment,length))
  logRR <- vector(length=trt.no)
  seRR <- vector(length=trt.no)
  
  for(i in 1:trt.no){
    logRR[i] <- round(log(means[i+1])-log(means[1]),3)
    sp2 <- (((ns[1]-1)*sds[1]^2+(ns[i+1]-1)*sds[i+1]^2)/(ns[1]+ns[i+1]-2))
    seRR[i] <- round(sqrt(sp2*(1/(means[1]^2*ns[1])+1/(means[i+1]^2*ns[i+1]))),3)
    
  }
  res <- data.frame(trt[-1],logRR,seRR,row.names=NULL)
  colnames(res)<-c("Treatment","logRR","seRR")
  return(res)
}

```


##Data prep
```{r}
#HAB2017 <- read.csv("HAB2017_incub.csv")

#Resorts the treatments to a better order
HAB2017$Treatment <- factor(HAB2017$Treatment,levels=c("Initial","Control","NH4","PO4","Trace","N+P","N+trace","P+trace","N+P+trace"))
HAB2017$Nit <- factor(HAB2017$Nit,levels=c("NoN","NH4"))
HAB2017$Pho <- factor(HAB2017$Pho,levels=c("NoP","PO4"))
HAB2017$Trace <- factor(HAB2017$Trace,levels=c("NoT","Trace"))

#Chelex treatments had +10 mg/treatment. Correct all values by -20 mg/L
HAB2017[HAB2017$Treatment=="Chelex",]$Dry_mass <- HAB2017[HAB2017$Treatment=="Chelex",]$Dry_mass-20

summary(HAB2017)

#MC data
MC <- read.csv("~/Documents/GitHub/algalanalysis/Daves code/incub_MC.csv")
MCag <- aggregate(MC$MC,by=list(MC$Treatment,MC$Site),mean)
colnames(MCag)[3]<-"MC"

#Merge data
HAB2017$TS <- paste(HAB2017$Site,HAB2017$Treatment)
MCag$TS <- paste(MCag$Group.2,MCag$Group.1)
inc2 <- merge(HAB2017,MCag[,3:4],by="TS")

#Site subsets
Ford <- subset(inc2,inc2$Site=="Ford")
GB <- subset(inc2,inc2$Site=="GB")
NEr <- subset(inc2,inc2$Site=="NErie")
Mau <- subset(inc2,inc2$Site=="Maumee")

Inc.trim <- subset(inc2,inc2$Treatment!="Initial"&inc2$Treatment!="Chelex")
Inc.trim$Treatment <- factor(Inc.trim$Treatment)[drop=T,]

Ford.trim <- subset(Inc.trim,Site=="Ford")
Mau.trim <- subset(Inc.trim,Site=="Maumee")
NEr.trim <- subset(Inc.trim,Site=="NErie")
GB.trim <- subset(Inc.trim,Site=="GB")

```

##Calculate log response ratios
```{r}
FordRR <- logRR(Ford.trim)
MauRR <- logRR(Mau.trim)
NErRR <- logRR(NEr.trim)
```


##Incubator experiment - dry mass
Data exploration plots

```{r}
plot(Dry_mass~Treatment,data=Ford,las=2,ylab="Dry mass (mg/L)")
plot(Dry_mass~Treatment,data=GB,las=2,ylab="Dry mass (mg/L)")
```

##Incubator experiment - chlorophyll a
Data exploration plots looking at log response ratios
These plots were used for presentations at WMAO and Slippery Rock (2017/2018)

```{r}
plot(Chla*1000~Treatment,data=Ford,las=2,ylab="Chlorophyll a (µg/L)")
plot(Chla*1000~Treatment,data=GB,las=2,ylab="Chlorophyll a (µg/L)")
plot(Chla*1000~Treatment,data=Mau,las=2,ylab="Chlorophyll a (µg/L)", main="Maumee Chlorophyll-a Response")
plot(Chla*1000~Treatment,data=NEr,las=2,ylab="Chlorophyll a (µg/L)", main="North Erie Chlorophyll-a Response")

#Vector of colors for treatments
cs <- c("orangered3","seagreen4","black","sienna2","chocolate4","coral3","dodgerblue")
####Where does MauRR come from? 
MauRR<-MauRR[c(1,2,4,3,5,6,7,8),]

par(pin=c(3,2))
xloc <- barplot(MauRR$logRR[1:7],las=1,ylim=c(-1,1),ylab="ln(Treatment/Control)",xaxt="n",border=NA,col=cs)
with(MauRR[1:7,],arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
mtext(MauRR$Treatment[1:7],side=1,at=xloc,las=2,line=1)
abline(h=c(0.385,-0.385,0),lwd=2,lty=c(2,2,1))
box()

NErRR<-NErRR[c(1,2,4,3,5,6,7,8),]

par(pin=c(3,2))
xloc <- barplot(NErRR$logRR[1:7],las=1,ylim=c(-1,3.5),ylab="ln(Treatment/Control)",xaxt="n",border=NA,col=cs)
with(NErRR[1:7,],arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
mtext(NErRR$Treatment[1:7],side=1,at=xloc,las=2,line=1)
abline(h=c(0.385,-0.385,0),lwd=2,lty=c(2,2,1))
box()
```


##Incubator MC concentration
Plots of raw chlorophyll a concentration data in all treatments (i.e., no log RR). Plots also included an overlay of microcystin concentrations (data from LH Larson @ USGS). Plots were used for project update call with James Larson (Mar 2018).

```{r}
HAB2017$Treatment <- factor(HAB2017$Treatment,levels=c("Initial","Control","NH4","PO4","Trace","N+P","N+trace","P+trace","N+P+trace"))
HAB2017$Nit <- factor(HAB2017$Nit,levels=c("NoN","NH4"))
HAB2017$Pho <- factor(HAB2017$Pho,levels=c("NoP","PO4"))
HAB2017$Trace <- factor(HAB2017$Trace,levels=c("NoT","Trace"))
#Maumee
par(mfrow=c(1,1)) 
par(pin=c(4,3),mar=c(6,4,1,4))
plot(Chla*1000~Treatment,data=Mau,las=2,xlab="Treatment",ylim=c(0,250),ylab="Chlorophyll a (ug/L)")
stripchart(MC*25~Treatment,data=Mau,vertical=T,add=T,cex=2,col="red",lwd=2,pch=18)
axis(4,las=2,labels=seq(0,10,by=2),at=seq(0,250,by=50),col="red",col.axis="red")
mtext("Microcystin (ug/L)",4,line=2,col="red")
legend("top","Maumee",bty="n",cex=1.4)
#Green Bay
par(pin=c(4,3))
plot(Chla*1000~Treatment,data=GB,las=2,xlab="",ylim=c(0,250),ylab="Chlorophyll a (ug/L)")
stripchart(MC*25~Treatment,data=GB,vertical=T,add=T,cex=2,col="red",lwd=2,pch=18)
axis(4,las=2,labels=seq(0,10,by=2),at=seq(0,250,by=50),col="red",col.axis="red")
mtext("Microcystin (ug/L)",4,line=2,col="red")
legend("top","Green Bay",bty="n",cex=1.4)
#North Erie
par(mfrow=c(1,1)) 
par(pin=c(4,3),mar=c(6,4,1,4))
plot(Chla*1000~Treatment,data=NEr,las=2,xlab="",ylim=c(0,250),ylab="Chlorophyll a (ug/L)")
stripchart(MC*25~Treatment,data=NEr,vertical=T,add=T,cex=2,col="red",lwd=2,pch=18)
axis(4,las=2,labels=seq(0,3,by=0.5),at=seq(0,250,by=50),col="red",col.axis="red")
mtext("Microcystin (ug/L)",4,line=2,col="red")
legend("top","North Erie",bty="n",cex=1.4)

#Ford
par(pin=c(4,3))
plot(Chla*1000~Treatment,data=Ford,las=2,xlab="",ylim=c(0,250),ylab="Chlorophyll a (ug/L)")
stripchart(MC*25~Treatment,data=Ford,vertical=T,add=T,cex=2,col="red",lwd=2,pch=18)
axis(4,las=2,labels=seq(0,10,by=2),at=seq(0,250,by=50),col="red",col.axis="red")
mtext("Microcystin (ug/L)",4,line=2,col="red")
legend("top","Ford",bty="n",cex=1.4)

#Ford with smaller y-axis scales
par(pin=c(4,3))
plot(Chla*1000~Treatment,data=Ford,las=2,xlab="",ylim=c(0,50),ylab="Chlorophyll a (ug/L)")
stripchart(MC*25~Treatment,data=Ford,vertical=T,add=T,cex=2,col="red",lwd=2,pch=18)
axis(4,las=2,labels=seq(0,2,by=0.4),at=seq(0,50,by=10),col="red",col.axis="red")
mtext("Microcystin (ug/L)",4,line=2,col="red")
legend("top","Ford",bty="n",cex=1.4)
abline(h=0.1*25,col="red",lty=2)

#N Erie with smaller y-axis scales
par(pin=c(4,3))
par(pin=c(4,3),mar=c(6,4,1,4))
plot(Chla*1000~Treatment,data=NEr,las=2,xlab="Treatment",ylim=c(0,120),ylab="Chlorophyll a (ug/L)")
stripchart(MC*60~Treatment,data=NEr,vertical=T,add=T,cex=2,col="red",lwd=2,pch=18)
axis(4,las=2,labels=seq(0,2,by=0.4),at=seq(0,120,by=24),col="red",col.axis="red")
mtext("Microcystin (ug/L)",4,line=2,col="red")
legend("top","North Erie",bty="n",cex=1.4)

```

##Incubator stats
###Maumee
```{r Maumee}
plot(Chla~Treatment,data=Mau,las=2)
MI.ch <- aov(Chla~Nit*Pho*Trace,data=Mau.trim)
summary(MI.ch)
summary(aov(Chla~Nit+Pho+Trace+Nit:Trace,data=Mau.trim))
shapiro.test(resid(MI.ch))
plot(MI.ch,which=1)

plot(AFDM~Treatment,data=Mau,las=2)
MI.dm <- aov(AFDM~Nit*Pho*Trace,data=Mau.trim)
summary(MI.dm)
summary.lm(MI.dm)
shapiro.test(resid(MI.dm))
plot(MI.dm,which=1)

plot(MC~as.numeric(Treatment),data=Mau.trim,las=2,xaxt="n",xlab="")
axis(1,at=seq(1,8),labels=levels(Mau.trim$Treatment),las=2)
```

Chla was limited by N and P and weakly by N+trace (p = 0.075). AFDM also responded to P but with a negative interaction between P+trace. Microcystin was elevated in N+trace and less so in N+P+trace.

###North Erie
Should do a log transformation but there are some neg values.
```{r N. Erie}
plot(Chla~Treatment,data=NEr,las=2)
NI.ch <- aov(Chla~Nit*Pho*Trace,data=NEr.trim)
summary(NI.ch)
shapiro.test(resid(NI.ch))
plot(NI.ch,which=1)

plot(AFDM~Treatment,data=NEr,las=2)
NI.dm <- aov(AFDM~Nit*Pho*Trace,data=NEr.trim)
summary(NI.dm)
summary(lm(AFDM~Pho*Trace+Nit+Pho:Nit,data=NEr.trim))
shapiro.test(resid(NI.dm))
plot(NI.dm,which=1)

plot(MC~as.numeric(Treatment),data=NEr.trim,las=2,xaxt="n",xlab="")
axis(1,at=seq(1,8),labels=levels(NEr.trim$Treatment),las=2)
```

Chla was primary P, secondary N, and tertiary trace metal limited. For AFDM, NxP and TracexP stimulated growth but no elements alone. Trace+P response was stronger than N+P. P alone had an increase in MC but P with trace had lower MC concentrations.

###Green Bay
```{r Green Bay}
plot(Chla~Treatment,data=GB,las=2)
GI.ch <- aov(Chla~Nit*Pho*Trace,data=GB.trim)
summary(GI.ch)
shapiro.test(resid(GI.ch))
plot(GI.ch,which=1)

plot(AFDM~Treatment,data=GB,las=2)
GI.dm <- aov(AFDM~Nit*Pho*Trace,data=GB.trim)
summary(GI.dm)
summary(lm(AFDM~Nit*Pho,data=GB.trim))
shapiro.test(resid(GI.dm))
plot(GI.dm,which=1)

plot(MC~as.numeric(Treatment),data=GB.trim,las=2,xaxt="n",xlab="")
axis(1,at=seq(1,8),labels=levels(GB.trim$Treatment),las=2)
```

Strong N limitation for both Chla and AFDM. Slight inhibition in AFDM of N+P treatments. Adding N also stimulated MC production.

###Ford
Also needs a log transformation but some neg values of chla
```{r Green Bay}
plot(Chla~Treatment,data=Ford,las=2)
FI.ch <- aov(Chla~Nit*Pho*Trace,data=Ford.trim)
summary(FI.ch)
shapiro.test(resid(FI.ch))
plot(FI.ch,which=1)

plot(AFDM~Treatment,data=Ford,las=2)
FI.dm <- aov(AFDM~Nit*Pho*Trace,data=Ford.trim)
summary(FI.dm)
shapiro.test(resid(FI.dm))
plot(FI.dm,which=1)

plot(MC~as.numeric(Treatment),data=Ford.trim,las=2,xaxt="n",xlab="")
axis(1,at=seq(1,8),labels=levels(Ford.trim$Treatment),las=2)
```

P limitation for chla and AFDM (stronger for chla), weak N (p=0.055) and N+P (p=0.087) for chla. All MC below DL.


###MC across all sites
```{r MC data}

plot(MC~as.numeric(Treatment),data=Inc.trim,xaxt="n",xlab="",log="y",las=2)
axis(1,at=seq(1,8),labels=levels(Inc.trim$Treatment),las=2)

MCinc <- lmer(log(MC)~Nit*Pho*Trace+(1|Site),data=Inc.trim,na.action=na.omit)
summary(MCinc)
plot(MCinc)

MCinc2 <- update(MCinc,.~.-Nit:Pho:Trace)
anova(MCinc2,MCinc)
#Three-way interaction is NS

MCinc3 <- update(MCinc2,.~.-Pho:Trace)
anova(MCinc3,MCinc2)
#P:Trace is sig p=0.012

MCinc4 <- update(MCinc2,.~.-Pho:Nit)
anova(MCinc4,MCinc2)
#N:P is NS

MCinc5 <- update(MCinc4,.~.-Nit:Trace)
anova(MCinc5,MCinc4)
#N:Trace is NS

MCinc6 <- update(MCinc5,.~.-Nit)
anova(MCinc6,MCinc5)
#N is significant (p < 0.001)

summary(MCinc5)
r.squaredGLMM(MCinc5)
```

Over every all sites, P+trace and nitrogen are influencing MC production. 99% of variance explained by site-to-site differences with just 0.3% related to fixed treatments.
