---
title: "NDS USGS/KSU HABS 2017"
author: "Dave Costello"
date: "10/2/2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/dcostel3/Dropbox/ASAP/2017 USGS/algalanalysis/Daves code/")
#setwd("C:/Users/dcostel3/Dropbox/ASAP/2017 USGS/algalanalysis/Daves code/")

options(device=quartz)
library(nlme)
library(lme4)
library(MuMIn)
```

```{r log response ratio function for chlorophyll a}
logRR <- function(dataset){
  dataset$TRT <- dataset$TRT[drop=T]
  trt <- levels(dataset$TRT)
  trt.no <- length(trt)-1
  means <- with(dataset,tapply(CHLA,TRT,mean,na.rm=T))
  sds <- with(dataset,tapply(CHLA,TRT,sd,na.rm=T))
  ns <- with(dataset,tapply(CHLA,TRT,length))
  logRR <- vector(length=trt.no)
  seRR <- vector(length=trt.no)
  
  for(i in 1:trt.no){
    logRR[i] <- round(log(means[i+1])-log(means[1]),3)
    sp2 <- (((ns[1]-1)*sds[1]^2+(ns[i+1]-1)*sds[i+1]^2)/(ns[1]+ns[i+1]-2))
    seRR[i] <- round(sqrt(sp2*(1/(means[1]^2*ns[1])+1/(means[i+1]^2*ns[i+1]))),3)
    
  }
  res <- data.frame(trt[-1],logRR,seRR,row.names=NULL)
  colnames(res)<-c("Treatment","logRR","seRR")
  return(res)
}

```


```{r import data}
NDS <- read.csv(file="NDS_chla.csv")
NDS$TRT <- factor(NDS$TRT,levels=c("Control","NH4","PO4","Fe","Zn","Ni&Mo",
                                   "N&Zn","N&Fe","P&Ni&Mo","N&P"))
NDS$N <- factor(NDS$N,levels=c("noN","NH4"))
NDS$Trace <- factor(NDS$Trace,levels=c("noT","Fe","NiMo","Zn"))

MC <- read.csv(file="NDS_MC.csv")
MC$TRT <- factor(MC$TRT,levels=c("Control","NH4","PO4","Fe","Zn","Ni&Mo",
                                   "N&Zn","N&Fe","P&Ni&Mo","N&P"))
MC$N <- factor(MC$N,levels=c("noN","NH4"))
MC$Trace <- factor(MC$Trace,levels=c("noT","Fe","NiMo","Zn"))
summary(MC)

#Separate by sites and calculate log response ratios
OWC <- subset(NDS,NDS$SITE=="OLD WOMAN 1")
SAN <- subset(NDS,NDS$SITE=="SANDUSKY 1")
MA1 <- subset(NDS,NDS$SITE=="MAUMEE 1")
MA2 <- subset(NDS,NDS$SITE=="MAUMEE 2")
NE1 <- subset(NDS,NDS$SITE=="N ERIE 1")
NE2 <- subset(NDS,NDS$SITE=="N ERIE 2")
FD1 <- subset(NDS,NDS$SITE=="FORD 1")
FD2 <- subset(NDS,NDS$SITE=="FORD 2")
GB <- subset(NDS,NDS$SITE=="GREEN BAY 1")
FOX <- subset(NDS,NDS$SITE=="FOX 1")

#Outliers - check raw data
NE2[NE2$AFDM>2,'AFDM'] <-NA
#Negative CHLA in SAN CTRL
```

```{r calc. log RR}
OWCRR <- logRR(OWC)
SANRR <- logRR(SAN)
MA1RR <- logRR(MA1)
MA2RR <- logRR(MA2)
NE1RR <- logRR(NE1)
NE2RR <- logRR(NE2)
FD1RR <- logRR(FD1)
FD2RR <- logRR(FD2)
GBRR <- logRR(GB)
FOXRR <- logRR(FOX)
ALLRR <- logRR(NDS)
```


```{r exploratory plots}
plot(CHLA~TRT,data=NDS,las=2)
plot(CHLA~SITE,data=NDS,las=2,xlab="",ylab="Chla (mg/m2)")
plot(CHLA~TRT,data=FD2)

site="FOX 1"
plot(CHLA~TRT,ylab="Chla (mg/m2)",las=2,data=NDS[NDS$SITE==site,],
     ylim=c(0,300),main=site,xlab="")

#Log RR plots
#Old Woman
xloc <- with(OWCRR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Old Woman"))
with(OWCRR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#Sandusky
xloc <- with(SANRR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Sandusky"))
with(SANRR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#Maumee 1
xloc <- with(MA1RR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Maumee 1"))
with(MA1RR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#Maumee 2
xloc <- with(MA2RR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Maumee 2"))
with(MA2RR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#N. Erie 1
xloc <- with(NE1RR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="N. Erie 1"))
with(NE1RR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#N. Erie 2
xloc <- with(NE2RR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="N. Erie 2"))
with(NE2RR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#Ford 1
xloc <- with(FD1RR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Ford 1"))
with(FD1RR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#Ford 2
xloc <- with(FD2RR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Ford 2"))
with(FD2RR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#Green Bay
xloc <- with(GBRR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Green Bay"))
with(GBRR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#Fox
xloc <- with(FOXRR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="Fox"))
with(FOXRR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

#ALL
xloc <- with(ALLRR,
             barplot(logRR,names.arg=Treatment,las=2,ylim=c(-1,3),
             ylab="log response ratio (Treatment/Control)",
             main="ALL"))
with(ALLRR,
     arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
abline(h=c(-0.385,0.385,0),lwd=2,lty=c(2,2,1))
box()

```

```{r presentation plots}
MA2RR<-MA2RR[c(1,2,9,3,4,5,6,7,8),]
NE1RR<-NE1RR[c(1,2,9,3,4,5,6,7,8),]

cs <- c("orangered3","seagreen4","black","sienna2","orchid4","lightcoral",
        "chocolate4","purple4","coral3")



par(pin=c(3,2))
xloc <- barplot(MA2RR$logRR,las=1,ylim=c(-1,1),ylab="ln(Treatment/Control)",xaxt="n",border=NA,col=cs)
with(MA2RR,arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
mtext(MA2RR$Treatment,side=1,at=xloc,las=2,line=1)
abline(h=c(0.385,-0.385,0),lwd=2,lty=c(2,2,1))
box()

xloc <- barplot(NE1RR$logRR,las=1,ylim=c(-1,2),ylab="ln(Treatment/Control)",xaxt="n",border=NA,col=cs)
with(NE1RR,arrows(xloc,(logRR+seRR),xloc,(logRR-seRR),length=0.05,angle=90,code=3))
mtext(NE1RR$Treatment,side=1,at=xloc,las=2,line=1)
abline(h=c(0.385,-0.385,0),lwd=2,lty=c(2,2,1))
box()

```

###Site maps

```{r site maps}
library(maps)
library(mapdata)
lkNDS <- read.csv(file="lakeNDSsites.csv")

#BIG MAP
pdf(file="fullmap.pdf",height=4,width=4)

map("worldHires","US",xlim=c(-89,-81),ylim=c(41,47),col="white")
map("state",c("michigan","ohio","indiana","pennsylvania","illinois","west virginia","wisconsin"),add=T,col="grey90",fill=T)
map("worldHires","Canada",add=T,col="grey90",fill=T)
rect(-83.7,41.2,-82.2,42.2,lwd=1.2,border="red")
rect(-88.3,44.4,-86.5,45.9,lwd=1.2,border="red")
text(-83.7,42.2,labels="6",adj=c(-0.5,1.5),col="red",cex=0.5)
text(-88.3,45.9,labels="4",adj=c(-5,1.5),col="red",cex=0.5)
points(lkNDS$Long,lkNDS$Lat,cex=0.5,pch=16,col="red")
box()

dev.off()

#LAKE ERIE MAP
pdf(file="OHmap.pdf",height=4,width=4)

map("state","ohio",xlim=c(-83.7,-82.2),ylim=c(41.2,42.2),col="grey90",fill=T)
map("state","michigan",xlim=c(-86,-80),ylim=c(40,45),add=T,col="grey90",fill=T)
map("worldHires","Canada",xlim=c(-86,-80),ylim=c(40,45),add=T,col="grey90",fill=T)
box()
points(lkNDS$Long,lkNDS$Lat,cex=1,col="red")
points(Lat~Long,data=lkNDS[lkNDS$Incub==T,],pch=16,col="red",cex=1)
points(Lat~Long,data=lkNDS[lkNDS$Incub==T,],pch=1,col="black",cex=1)

#State names
text(-83.2,41.5,"OHIO")
text(-83.45,42.05,"MICHIGAN")
text(-82.7,42.1,"ONTARIO")

#Site names
text(-82.6,41.5,"Sandusky",cex=0.5)
text(-83.3,41.75,"Maumee",cex=0.5)
text(-83.1,41.9,"N. Erie",cex=0.5)
text(-82.35,41.35,"Old Woman",cex=0.5)

dev.off()

#LAKE MICHIGAN MAP
pdf(file="WImap.pdf",height=4,width=3.5)

map("state","wisconsin",xlim=c(-88.3,-86.5),ylim=c(44.4,45.9),col="grey90",fill=T)
map("state","michigan",xlim=c(-88.3,-86.5),ylim=c(44.4,45.9),col="grey90",fill=T,add=T)
box()

points(lkNDS$Long,lkNDS$Lat,cex=1,col="red")
points(Lat~Long,data=lkNDS[lkNDS$Incub==T,],pch=16,col="red",cex=1)
points(Lat~Long,data=lkNDS[lkNDS$Incub==T,],pch=1,col="black",cex=1)

#State names
text(-87.5,45.8,"MICHIGAN")
text(-88,45,"WISC.")

#Site names
text(-87.9,44.7,"Green Bay/Fox",cex=0.5)
text(-87.0,45.65,"Ford",cex=0.5)

dev.off()
```

```{r River map}
#DEFINE URL
# - this is the location of the file
url.river_data <- url("http://sharpsightlabs.com/wp-content/datasets/usa_rivers.RData")


# LOAD DATA
# - this will retrieve the data from the URL
load(url.river_data)

lines.rivers <- subset(lines.rivers, !(FEATURE %in% c("Shoreline"
                                                      ,"Shoreline Intermittent"
                                                      ,"Null"
                                                      ,"Closure Line"
                                                      ,"Apparent Limit"
                                                      )))
OHrivers <- subset(lines.rivers, (STATE %in% 'OH'))
MIrivers <- subset(lines.rivers, (STATE %in% 'MI'))


map("worldHires","US",xlim=c(-90,-80),ylim=c(40,48),col="white")
map("state",c("michigan","ohio","indiana","pennsylvania","illinois","west virginia","wisconsin"),add=T,col="grey90",fill=T)
map("worldHires","Canada",add=T,col="grey90",fill=T)

stNDS <- read.csv(file="streamNDSsites.csv")
points(-87.8,46.9,pch=22,cex=4,col="red")
points(-81.5,41.3,pch=22,cex=5,col="red")
points(stNDS$Long,stNDS$Lat,cex=1.2,col="red",pch=16)

map("state","ohio",xlim=c(-82,-81),ylim=c(41,42),fill=T,col="grey90")
plot(OHrivers,add=T)
points(stNDS$Long,stNDS$Lat,cex=1.2,col="red",pch=16)

map("state","michigan",xlim=c(-88,-87.75),ylim=c(46.7,47),fill=T,col="grey90")
plot(MIrivers,add=T)
points(stNDS$Long,stNDS$Lat,cex=1.2,col="red",pch=16)

```

##Linear models
Because the treatments are not fully crossed, design is non-orthogonal. Need to look at lm output not aov output for interpretation.

###Old Woman Creek
```{r OWC stats}
plot(CHLA~TRT,data=OWC,las=2)
OWC.ch <- lm(CHLA~N*P*Trace,data=OWC)
summary(OWC.ch)
plot(OWC.ch,which=1)
shapiro.test(resid(OWC.ch))

plot(AFDM~TRT,data=OWC,las=3)
OWC.dm <- lm(AFDM~N*P*Trace,data=OWC)
summary(OWC.dm)
shapiro.test(resid(OWC.dm))
plot(OWC.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="OLD WOMAN 1",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

N and P colimited (Chla) and N limited (AFDM). MC was only observed in 1 rep of N&P.

###Sandusky
```{r Sandusky stats}
plot(CHLA~TRT,data=SAN,las=2)
SAN.ch <- lm(CHLA~N*P*Trace,data=SAN)
summary(SAN.ch)
shapiro.test(resid(SAN.ch))
plot(SAN.ch,which=1)

plot(AFDM~TRT,data=SAN,las=2)
SAN.dm <- lm(AFDM~N*P*Trace,data=SAN)
summary(SAN.dm)
shapiro.test(resid(SAN.dm))
plot(SAN.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="SANDUSKY 1",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

Strong P limitation, weak secondary N and Fe (Chla). No limitation for AFDM. AFDM in Sandusky was very high across the board, so nutrients just seemed to shift the autotrophic index to more chla rich biomass. Higher MC in Zn-only treatment and P+Ni+Mo(1 rep).


###North Erie 1
```{r N Erie 1 stats}
plot(CHLA~TRT,data=NE1,las=2)
NE1.ch <- lm(CHLA~N*P*Trace,data=NE1)
summary(NE1.ch)
shapiro.test(resid(NE1.ch))
plot(NE1.ch,which=1)

plot(AFDM~TRT,data=NE1,las=2)
NE1.dm <- lm(AFDM~N*P*Trace,data=NE1)
summary(NE1.dm)
shapiro.test(resid(NE1.dm))
plot(NE1.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="N. ERIE 1",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

Strong P effect (chla) and slight inhibition with N alone (AFDM) but the inhibition goes away when interactions are removed. No difference in AFDM. MC was observed in Fe (1 rep) and N&P (1 rep) treatments.

###North Erie 2
```{r N Erie 2 stats}
plot(CHLA~TRT,data=NE2,las=2)
NE2.ch <- lm(CHLA~N*P*Trace,data=NE2)
summary(NE2.ch)
shapiro.test(resid(NE2.ch))
plot(NE2.ch,which=1)

plot(AFDM~TRT,data=NE2,las=2)
NE2.dm <- lm(AFDM~N*P*Trace,data=NE2)
summary(NE2.dm)
shapiro.test(resid(NE2.dm))
plot(NE2.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="N. ERIE 2",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

No limitation. Very low biomass overall. Messy MC data, but some evidence of increase in NH4, N&Fe (1 rep), and P&Ni&Mo.

###Maumee
```{r Maumee grouped stats}
#Make a grouped Mamuee response
MAU <- rbind(MA1,MA2)

plot(CHLA~TRT,data=MAU,las=2,log="y")
MAU.ch <- lm(log(CHLA)~N*P*Trace,data=MAU)
summary(MAU.ch)
shapiro.test(resid(MAU.ch))
plot(MAU.ch,which=1)

plot(AFDM~TRT,data=MAU,las=2)
MAU.dm <- lm(AFDM~N*P*Trace,data=MAU)
summary(MAU.dm)
shapiro.test(resid(MAU.dm))
plot(MAU.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="MAUMEE 2",],las=2,xaxt="n",xlab="",ylim=c(0,0.12))
points(MC~as.numeric(TRT),data=MC[MC$SITE=="MAUMEE 1",],pch=16)
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)

MAUMC <- subset(MC,SITE=="MAUMEE 2"|SITE=="MAUMEE 1")
MAU.mc <- lm(log(MC)~N*P*Trace,data=MAUMC)
summary(MAU.mc)
plot(MAU.mc)
```

Patterns are not statistically strong, but evidence that "single" element additions (N, P, Ni+Mo) reduce chla relative to control but mixtures of elements (P+Ni+Mo) can sustain chla. For AFDM P and Fe alone reduce biomass and N+Fe sustains it. MC data is messy but Zn seems to stimulate MC production. Also increase in P+Ni+Mo.

###Ford 1
```{r Ford 1}
plot(CHLA~TRT,data=FD1,las=2)
FD1.ch <- lm(CHLA~N*P*Trace,data=FD1)
summary(FD1.ch)
shapiro.test(resid(FD1.ch))
plot(FD1.ch,which=1)

plot(AFDM~TRT,data=FD1,las=2)
FD1.dm <- lm(AFDM~N*P*Trace,data=FD1)
summary(FD1.dm)
shapiro.test(resid(FD1.dm))
plot(FD1.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="FORD 1",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

Strong N:P co-limitation (chla). No response in AFDM. One rep of P+Ni+Mo produced microcystin.


###Ford 2
```{r Ford 2}
plot(CHLA~TRT,data=FD2,las=2,log="y")
FD2.ch <- lm(log(CHLA)~N*P*Trace,data=FD2)
summary(FD2.ch)
shapiro.test(resid(FD2.ch))
plot(FD2.ch,which=1)

plot(AFDM~TRT,data=FD2,las=2)
FD2.dm <- lm(AFDM~N*P*Trace,data=FD2)
summary(FD2.dm)
shapiro.test(resid(FD2.dm))
plot(FD2.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="FORD 2",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

Strong P limitation (chla). No response in AFDM. One rep each of P+Ni+Mo, N, and P produced microcystin.

###Fox River
```{r Fox}
plot(CHLA~TRT,data=FOX,las=2)
FOX.ch <- lm(CHLA~N*P*Trace,data=FOX)
summary(FOX.ch)
shapiro.test(resid(FOX.ch))
plot(FOX.ch,which=1)

plot(AFDM~TRT,data=FOX,las=2)
FOX.dm <- lm(AFDM~N*P*Trace,data=FOX)
summary(FOX.dm)
shapiro.test(resid(FOX.dm))
plot(FOX.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="FOX 1",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

Strong P limitation (chla). No response in AFDM. MC doesn't show any strong patterns (relatively high ctrls).


###Green Bay
```{r Green Bay}
plot(CHLA~TRT,data=GB,las=2)
GB.ch <- lm(CHLA~N*P*Trace,data=GB)
summary(GB.ch)
shapiro.test(resid(GB.ch))
plot(GB.ch,which=1)

plot(AFDM~TRT,data=GB,las=2)
GB.dm <- lm(AFDM~N*P*Trace,data=GB)
summary(GB.dm)
shapiro.test(resid(GB.dm))
plot(GB.dm,which=1)

plot(MC~as.numeric(TRT),data=MC[MC$SITE=="GREEN BAY 1",],las=2,xaxt="n",xlab="")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)
```

Zn limitation of AFDM but additional N counteracted that increase. Patterns are the same in chla but not significant. MC was stimulated in one rep each by Ni+Mo, P+Ni+Mo, N+Zn, N+Fe.


###MC across all sites
```{r MC data}
library(lme4)

MC[MC$SITE=="MAUMEE 1",'SITE'] <- "MAUMEE 2"
plot(MC~as.numeric(TRT),data=MC,xaxt="n",xlab="",log="y")
axis(1,at=seq(1,10),labels=levels(MC$TRT),las=2)

MCmod <- lmer(log(MC)~N+P+Trace+N:P+N:Trace+P:Trace+(1|SITE),
              data=MC,na.action=na.omit)
summary(MCmod)
plot(MCmod)

MCmod2 <- update(MCmod,.~.-P:Trace)
anova(MCmod2,MCmod)

MCmod3 <- update(MCmod,.~.-N:Trace)
anova(MCmod3,MCmod)

MCmod4 <- update(MCmod3,.~.-N:P)
anova(MCmod4,MCmod3)

MCmod5 <- update(MCmod4,.~.-N)
anova(MCmod5,MCmod4)

summary(MCmod5)
r.squaredGLMM(MCmod5)
```

Over every all sites, P+Ni+Mo is the only combination that significantly increases MC above controls (p = 0.002). Fixed factors explain 1.4% of variation and random effect explains 80% of variation.
